`Big data' og politologisk datavidenskab

### 2020-03-26 draft_v01.Rmd
start research design section

Big data in political science research designs:
  - addressing confounding
    - it's dirty (re: Salganik) -- human behavior is mixed together with actions taken by bots/automated systems
    - drifting -- Needs of *actual* system maintainers / users may be completely orthogonal to needs of political science research (or even opposed, consider FB)
    - algorithmically confounded -- large-scale systems have robot nannies. These include everything from spell checkers to YouTube's recommendation algorithm. The observed behavior we find in big data is a consequence of the (generally) unobservable interaction between humans and these algorithms.
    - Standard sampling issues (nonrepresentative, systematic sampling bias)
  - its role
    - uncommon to directly analyze
    - more often: measurement

>>>>

Big data has applications in political science in empirical studies of all types, from description to explanation, prediction, and causal studies. Although these applications are only beginning across the social sciences, the past 15 years have provided enough experience that we can already point to one strong finding that can always guide us when we apply big data in our work: research design still matters [see @toshkov16, p. 13; @CG15].

*Familiar sampling bias*

Perhaps the most heady idea about big data was expressed best in the, now infamous, article in *Wired* magazine by Chris @anderson08 -- the idea that big data eliminates sampling problems because *n*=all. That is, one can analyze all of the data. The article has met with a good deal of attention and pushback in the years since it was published. In early 2020, it boasts more than 2.000 citations on Google Scholar. To put that in perspective, Maurice Duverger's book *Political Parties*, the origin of political science's only finding that rises to the level of a law, was published in 1959 and counts just over 7.000 citations.

The concept of "*n*-all" implies that big data can be taken at face value. Its patterns reveal a complete picture of human behavior. Unfortunately, the selection problem is still with us. It arises whenever observations enter your data for systematic reasons. @AP08 illustrate the problem with an example from the National Health Interview Survey in the US, in which individuals are asked to rate their health on a scale. Perhaps surprisingly, they find that individuals in hospitals systematically rate their health as worse than those out of hospital. Is the obvious conclusion that hospitals make people sicker? Of course not, individuals in hospital *selected* into being in the hospital because they were sicker.

Similar situations often arise with big data. For example, @harford14 recounts the story of Boston's experiment with using cell phone tracking to identify pot holes in the streets that need repair. The result was that the city discovered every pothole in neighborhoods frequented by young, affluent drivers--they type of person who owned a smart phone and downloaded the city's app. Likewise, @TSSW10 confirm the finding in previous research that Twitter mentions of German parties correlated strongly with their vote share in the 2009 parliamentary elections, with one major exception. The extremely online *Pirate Party* garned a huge number of mentions on Twitter, while receiving a tiny fraction of the actual vote. In both of these cases, if *n*=all then it equals "all" of a specific, self-selected group of technology users.

Ignoring sampling bias has always meant that researchers risked finding the wrong answer. Using big data with biased samples simply means now we get extremely precise estimates of the wrong answer.

*Special headaches for big data*

Big data also confronts us with new sources of bias and confounding. Perhaps the most famous example for social scientists is that of Google Flu Trends (GFT). GFT was a Google project launched in 2008 that predicted regional flu epidemics from fine-grained data on users queries to Google about their symptoms [@Getal2009]. After initially receiving attention for its impressive accuracy, GFT's predictive performance began to diminish systematically over time until--by the time the project was shuttered in 2015--for years it had produced inaccurately high estimates sometimes as great as double the data reported by US Centers for Disease Control and Prevention [@harford14]. The reasons for GFT's collapse are instructive for understanding both why research design still matters and what new headaches come with big data.

Several things contributed to GFT's problems, but the most noteworth seems to have arises after Google adjusted its main search service [see @LKKV14]. Google, like many organizations that produce and process big data, is constantly updating its service. In 2011 and 2012, Google added functions that suggested additional search terms to users (based on their initial search terms) and even suggested possible diagnoses for search terms that involved symptoms. The upshot was a feedback loop: users of Google search were nudged to refine their searches for symptoms and to seek out information on the flu, intensifying the signals that GFT relied on to predict flu outbreaks. There is more to GFT's story, of course, but this highlights some important principles for researchers to keep in mind.

Big data in political science research designs:
  - addressing confounding
    - it's dirty (re: Salganik) -- human behavior is mixed together with actions taken by bots/automated systems
    - drifting -- Needs of *actual* system maintainers / users may be completely orthogonal to needs of political science research (or even opposed, consider FB)
    - algorithmically confounded -- large-scale systems have robot nannies. These include everything from spell checkers to YouTube's recommendation algorithm. The observed behavior we find in big data is a consequence of the (generally) unobservable interaction between humans and these algorithms.

## measurement stuff

<!-- here we transition to the discussion of units of analysis and jusifying the focus on measurement -->

- role of big data in the research design
    - uncommon to directly analyze
    - more often: measurement

### 2020-03-26 draft_v01.tex
start research design section

Studiet af Bond m.fl. er exceptionelt fordi det kombinerer kvaliteterne
ved big data og eksperimentel metode. Mange forskere gør derfor også en
stor indsats for at etablere samarbejder med virksomheder og
organisationer der kan give dem adgang til data der ellers ville være
utilgængelige. Men samarbejde med virksomheder om big data er ikke uden
faldgruber. For det første kræver det ofte et betydeligt bureaukratisk
benarbejde at etablere et samarbejde. For det andet, og mere principielt
problematisk, er virksomheder og organisationer sjældent interesserede i
forskning der stiller dem selv i et dårligt lys. Det kan betyde at nogle
typer undersøgelser prioriteres på bekostningen af andre alene fordi de
passer til store teknologivirksomheders dagsordener. Eksempelvis
konkluderede \citet{bond2012} at Facebook-kampagnen havde en gunstig
effekt på valgdeltagelse. Men det er uklart om forskerne havde haft
samme frihedsgrader til at studere de negative konsekvenser af at bruge
Facebook.

@@

Big data in political science research designs: - addressing confounding
- it's dirty (re: Salganik) -- human behavior is mixed together with
actions taken by bots/automated systems - drifting -- Needs of
\emph{actual} system maintainers / users may be completely orthogonal to
needs of political science research (or even opposed, consider FB) -
algorithmically confounded -- large-scale systems have robot nannies.
These include everything from spell checkers to YouTube's recommendation
algorithm. The observed behavior we find in big data is a consequence of
the (generally) unobservable interaction between humans and these
algorithms. - Standard sampling issues (nonrepresentative, systematic
sampling bias) - its role - uncommon to directly analyze - more often:
measurement

@@

\hypertarget{anskaffelse-af-data}{%
\subsection{Anskaffelse af data}\label{anskaffelse-af-data}}

\begin{itemize}
\tightlist
\item
  Getting data: scraping, APIs, text databases
\end{itemize}

\hypertarget{usuperviserede-tilgange}{%
\subsection{Usuperviserede tilgange}\label{usuperviserede-tilgange}}

\hypertarget{superviserede-tilgange}{%
\subsection{Superviserede tilgange}\label{superviserede-tilgange}}

\hypertarget{tekst-som-data}{%
\subsection{Tekst som data}\label{tekst-som-data}}

\begin{itemize}
\item
  Pattern discovery (dimensionality reduction -- a la argument in Lowe
  2013WP)

  \begin{itemize}
  \tightlist
  \item
    Generic data: Clustering, IRT, etc.
  \item
    Fundamental unity of goals and approach
  \item
    Variety in methods results from variation in:

    \begin{itemize}
    \tightlist
    \item
      Assumptions re: underlying model/geometry of latent space
    \item
      Related to above: something like, ``format'' of output
    \item
      Structure/nature of input data
    \item
      Amount of domain expertise applied to structure results
    \item
      Assumptions about what is correlated with what
    \item
      Level of computational intensity
    \end{itemize}
  \end{itemize}
\item
  Text: Topic modeling, text scaling, dictionaries
\item
  Classification / prediction
\item
  Explanatory modeling
\end{itemize}

>>>>

Studiet af Bond m.fl. er exceptionelt fordi det kombinerer kvaliteterne
ved big data og eksperimentel metode. Mange forskere gør derfor også en
stor indsats for at etablere samarbejder med virksomheder og
organisationer der kan give dem adgang til data, der ellers ville være
utilgængelige. Men samarbejde med virksomheder om big data er ikke uden
faldgruber. For det første kræver det ofte et betydeligt bureaukratisk
benarbejde at etablere et samarbejde. For det andet, og mere principielt
problematisk, er virksomheder og organisationer sjældent interesserede i
forskning der stiller dem selv i et dårligt lys. Det kan betyde at nogle
typer undersøgelser prioriteres på bekostning af andre, alene fordi de
passer bedre til store teknologivirksomheders dagsordener. Eksempelvis
konkluderede \citet{bond2012} at Facebook-kampagnen havde en gunstig
effekt på valgdeltagelse. Det er i sagens natur en flatterende
konklusion for Facebook. Men det er uklart om forskerne havde haft samme
frihedsgrader til at studere de negative konsekvenser af at bruge
Facebook.

@@

Big data has applications in political science in empirical studies of
all types, from description to explanation, prediction, and causal
studies. Although these applications are only beginning across the
social sciences, the past 15 years have provided enough experience that
we can already point to one strong finding that can always guide us when
we apply big data in our work: research design still matters
\citetext{\citealp[see][p.~13]{toshkov16}; \citealp{CG15}}.

\emph{Familiar sampling bias}

Perhaps the most heady idea about big data was expressed best in the,
now infamous, article in \emph{Wired} magazine by Chris
\citet{anderson08} -- the idea that big data eliminates sampling
problems because \emph{n}=all. That is, one can analyze all of the data.
The article has met with a good deal of attention and pushback in the
years since it was published. In early 2020, it boasts more than 2.000
citations on Google Scholar. To put that in perspective, Maurice
Duverger's book \emph{Political Parties}, the origin of political
science's only finding that rises to the level of a law, was published
in 1959 and counts just over 7.000 citations.

The concept of ``\emph{n}-all'' implies that big data can be taken at
face value. Its patterns reveal a complete picture of human behavior.
Unfortunately, the selection problem is still with us. It arises
whenever observations enter your data for systematic reasons.
\citet{AP08} illustrate the problem with an example from the National
Health Interview Survey in the US, in which individuals are asked to
rate their health on a scale. Perhaps surprisingly, they find that
individuals in hospitals systematically rate their health as worse than
those out of hospital. Is the obvious conclusion that hospitals make
people sicker? Of course not, individuals in hospital \emph{selected}
into being in the hospital because they were sicker.

Similar situations often arise with big data. For example,
\citet{harford14} recounts the story of Boston's experiment with using
cell phone tracking to identify pot holes in the streets that need
repair. The result was that the city discovered every pothole in
neighborhoods frequented by young, affluent drivers--they type of person
who owned a smart phone and downloaded the city's app. Likewise,
\citet{TSSW10} confirm the finding in previous research that Twitter
mentions of German parties correlated strongly with their vote share in
the 2009 parliamentary elections, with one major exception. The
extremely online \emph{Pirate Party} garned a huge number of mentions on
Twitter, while receiving a tiny fraction of the actual vote. In both of
these cases, if \emph{n}=all then it equals ``all'' of a specific,
self-selected group of technology users.

Ignoring sampling bias has always meant that researchers risked finding
the wrong answer. Using big data with biased samples simply means now we
get extremely precise estimates of the wrong answer.

\emph{Special headaches for big data}

Big data also confronts us with new sources of bias and confounding.
Perhaps the most famous example for social scientists is that of Google
Flu Trends (GFT). GFT was a Google project launched in 2008 that
predicted regional flu epidemics from fine-grained data on users queries
to Google about their symptoms \citep{Getal2009}. After initially
receiving attention for its impressive accuracy, GFT's predictive
performance began to diminish systematically over time until--by the
time the project was shuttered in 2015--for years it had produced
inaccurately high estimates sometimes as great as double the data
reported by US Centers for Disease Control and Prevention
\citep{harford14}. The reasons for GFT's collapse are instructive for
understanding both why research design still matters and what new
headaches come with big data.

Several things contributed to GFT's problems, but the most noteworth
seems to have arises after Google adjusted its main search service
\citep[see][]{LKKV14}. Google, like many organizations that produce and
process big data, is constantly updating its service. In 2011 and 2012,
Google added functions that suggested additional search terms to users
(based on their initial search terms) and even suggested possible
diagnoses for search terms that involved symptoms. The upshot was a
feedback loop: users of Google search were nudged to refine their
searches for symptoms and to seek out information on the flu,
intensifying the signals that GFT relied on to predict flu outbreaks.
There is more to GFT's story, of course, but this highlights some
important principles for researchers to keep in mind.

Big data in political science research designs: - addressing confounding
- it's dirty (re: Salganik) -- human behavior is mixed together with
actions taken by bots/automated systems - drifting -- Needs of
\emph{actual} system maintainers / users may be completely orthogonal to
needs of political science research (or even opposed, consider FB) -
algorithmically confounded -- large-scale systems have robot nannies.
These include everything from spell checkers to YouTube's recommendation
algorithm. The observed behavior we find in big data is a consequence of
the (generally) unobservable interaction between humans and these
algorithms.

\hypertarget{measurement-stuff}{%
\subsection{measurement stuff}\label{measurement-stuff}}

\begin{itemize}
\tightlist
\item
  role of big data in the research design

  \begin{itemize}
  \tightlist
  \item
    uncommon to directly analyze
  \item
    more often: measurement
  \end{itemize}
\end{itemize}

@@

Big data præsenterer både computationelle og analytiske udfordringer -
vi fokuserer på analytiske

Fællestræk: højdimensionalitet

Fælles for metoder: dimensionalitetsreduktion

Klassifikation ctr. skalering

usuperviserede tilgange

superviserede tilgange
